---
title: "Stat 380 Final Project"
author: "Omer Canca"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

# Goal
We would like to see what variables correlate to highly rated hotels in Las Vegas. This will help us decide the important factors when booking a Hotel in Vegas

# Simple Linear Regression

### Clear Workspace and Load Libraries
```{r}
rm(list=ls())
```

```{r}
library(dplyr)
library(ggplot2)
library(corrplot)
library(leaps)
library(lattice)
library(caret)
```

### Data Inspection
```{r}
data = read.csv(file = "C:/users/omerc/Downloads/Vegas.csv", head = TRUE, sep=";")
head(data)
str(data)
```

### Coding Variables
Spa = 1 for yes; 0 for no
Gym = 1 for yes; 0 for no
Pool = 1 for yes; 0 for no
Casino = 1 for yes; 0 for no
Free.internet = 1 for yes; 0 for no
Tennis.court = 1 for yes; 0 for no

```{r}
data$Spa<-ifelse(data$Spa=="YES",1,0)
data$Gym<-ifelse(data$Gym=="YES",1,0)
data$Pool<-ifelse(data$Pool=="YES",1,0)
data$Casino<-ifelse(data$Casino=="YES",1,0)
data$Free.internet<-ifelse(data$Free.internet=="YES",1,0)
data$Tennis.court<-ifelse(data$Tennis.court=="YES",1,0)
df <- dplyr::select_if(data, is.numeric)
```

### Backward Selection
We want to view which variable correlates the most with the "Score" Variable. This will allow hotels in Vegas to see what their guests view as most important.

We did this using backward stepwise selection. This process begins with a model containing all the variables, then removes the least significant ones one after the other. Then it stops when there is one left.
```{r}
br <- regsubsets(Score~., data = df, nvmax = 1,
                     method = "backward")
summary(br)
```
Our results tell us that Pool is the most significant variable.

### Building our SLR Model and Plotting
Using Pool to predict the score

We see that the variable Pool is significant.
```{r}
model1 = lm(Score ~ Pool, data = df)
summary(model1)
ggplot(df, aes(x=Pool, y=Score)) +
  geom_boxplot(fill='steelblue') +  facet_wrap(~ Pool) + ylim(0,5.5)
```

# Multiple Linear Regression
Now, we would like to see if we are better off using a model with more predictors

### Stepwise Regression
We will build models with 2, 3, 4, and 5 variables based on forward regression.

```{r}
fr <- regsubsets(Score~., data = df, nvmax = 5,
                     method = "forward")
summary(fr)
```
Our analysis tells us that our new models will be
```{r}
model2 = lm(Score ~ Pool + Free.internet, data = df)
model3 = lm(Score ~ Pool + Free.internet + Member.years, data = df)
model4 = lm(Score ~ Pool + Free.internet + Member.years + Gym, data = df)
model5 = lm(Score ~ Pool + Free.internet + Member.years+ Gym + Spa, data = df)
```

### AIC
We will use AIC to determine which model is the best of the three. AIC is a score that is used to determine which model is best based on prediction error. A lower AIC is better
```{r}
AIC(model2)
AIC(model3)
AIC(model4)
AIC(model5)
```

### ANOVA
We see that model 2 has the lowest AIC. We will compare to our SLR model to see which model is the best predictor for score.
```{r}
anova(model2, model1)
```
Our results tell us that the extra predictor in model2 is significant.

### Final Model
We see that both the predictors in the model are significant. With our boxplots, we also see that hotels with a pool and free internet generally tend to have a higher score. This supports our model.
```{r}
summary(model2)
ggplot(df, aes(x=Pool, y=Score)) +
  geom_boxplot(fill='steelblue') +  facet_wrap(~ Pool) + ylim(0,5.5)
ggplot(df, aes(x=Free.internet, y=Score)) +
  geom_boxplot(fill='steelblue') +  facet_wrap(~ Pool) + ylim(0,5.5)
```


# Conclusion
Our final model includes free internet and pool as the best predictors for score. This tells us that when looking for a hotel in Vegas, we should look for these two predictors to find the hotels with the best experience. 
